Ribosome accumulation during early phase resistance training

### 2021-12-09 manuscript.Rmd
rewrite abstract

>>>>

**Aim**
To describe the time course of ribosome accumulation in response to resistance training with constant and variable volume and relate RNA accumulation to total UBF levels and muscle growth. 
**Methods**
Participants allocated to a training group (n=11) performed unilateral knee extension with constant (6 sets per session) and variable volume (6, 3, and 9 sets in sessions 1-4, 5-8, and 9-12, respectively) allocated to either leg. Muscle biopsies (m. vastus lateralis) were sampled in the training group before and 48 hrs after the first session and 48 hrs after sessions 4, 5, 8, 9, and 12, and after eight days of de-training. Muscle samples were also obtained from a non-training control group (n=7) at baseline and after 48 hrs and 3-5 weeks. Muscle specimens were assayed for total RNA, ribosomal RNA (rRNA), ribosomal protein S6 (rpS6), and upstream binding factor (UBF) protein content. Muscle thickness (ultrasound) and strength (isokinetic and isometric knee extension) were measured before and after resistance training or control periods.
**Results**
Resistance training led to muscle growth (0.84 mm, 95% CI: [0.05, 1.64] vs. control) with concomitant increases in total RNA, rRNA, UBF and rpS6 protein. Total RNA increased most rapidly in response to the first four sessions (8.6% [5.5, 11.7] per session), followed by a plateau and peak values (49.5% [34.2, 66.5] above baseline) after session 8. UBF protein levels were associated with total RNA levels after controlling for time, and increases in total RNA levels predicted RT-induced muscle hypertrophy. After de-training, total RNA and rRNA decreased without muscle mass changes indicating halted biosynthesis of ribosomes. In conclusion, ribosomes accumulate in the initial phase of RT and is sensitive to training cessation. The rate of accumulation predicts muscle hypertrophy
